---
title: "Wrangle OpenAQ extraordinary data"
author: M. Salmon
output: revealjs::revealjs_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE, 
                      cache = FALSE)
library("ggplot2")
library("ggmap")
library("viridis")
library("ggthemes")
library("ggrepel")
library("dplyr")
theme_set(theme_bw(base_size = 22))
```

# Intro: who am I

# Why wanting to get OpenAQ data?

# A treasure from New Zealand

## The All darkblues?

## The kiwis?

## No! R!

# Bringing OpenAQ to R!

## What is a package?

## `ropenaq`

## Quality control

# Some wrangling cases

## Map of all OpenAQ locations

## Delhi PM2.5 daily time series

```{r delhi hourly, fig.width = 10, fig.height = 6}
load("data/meas_india.RData")

table <- both %>% mutate(day = as.Date(dateLocal)) %>%
  filter(value > 0, value != 1985) %>%
  group_by(day, location) %>%
  summarize(PM25 = mean(value)) %>%
  ungroup()

 ggplot() +
  geom_line(data = table, aes(x = day,y = PM25), col = "darkblue")+
  ylab(expression(paste("PM2.5 concentration (", mu, "g/",m^3,")")))+
   xlab("Time (days)")
```

## vs. Indian health categories


```{r delhi daily, fig.width = 10, fig.height = 6}




paletteAQI <- viridis::viridis_pal(option = "plasma")(6)
paletteAQI2 <- factor(paletteAQI,
                     levels = paletteAQI)
labels <- factor(c( "Good", "Satisfactory", "Moderately polluted",
                    "Poor", "Very Poor", "Severe"),
                 levels = c( "Good", "Satisfactory", "Moderately polluted",
                             "Poor", "Very Poor", "Severe"))

aqiPM <- data.frame(ymin = c(0, 30, 60, 90, 120, 250),
                    ymax = c(30, 60, 90, 120, 250, Inf),
                    col = paletteAQI2)


ggplot() +
  geom_rect(data = aqiPM, aes(xmin = table$day[1], xmax = max(table$day),
                              ymin = ymin, ymax = ymax,
                              fill = col), alpha = 0.7)  +
  scale_fill_manual(name="AQI Category PM.5",
                      breaks=paletteAQI,
                      values=paletteAQI,
                      labels=labels)+
  geom_line(data = table, aes(x = day,y = PM25))+
  ylab(expression(paste("PM2.5 concentration (", mu, "g/",m^3,")")))+
   xlab("Time (days)") +
  theme(legend.position = "bottom")

```

## Delhi PM2.5 vs. WHO limit

```{r delhi daily who, fig.width = 10, fig.height = 6}

ggplot() +
  geom_line(data = table, aes(x = day,y = PM25), 
             col = "darkblue")+
  geom_hline(yintercept = 25, size = 1.2, col = "darkred") +
  ylab(expression(paste("PM2.5 concentration (", mu, "g/",m^3,")")))+
   xlab("Time (days)") 
```



## Sarajevo PM10 stations

```{r}
sarajevo <- readr::read_csv("data/sarajevo.csv")
daily <- filter(sarajevo, parameter == "pm10") %>%
  group_by(location, longitude, latitude, day = as.Date(dateLocal)) %>%
  summarize(pm10 = mean(value)) %>%
  group_by(day) %>%
  filter(all(!is.na(pm10))) %>%
  ungroup() %>%
  filter(location != "Ivan Sedlo")

map_data <- unique(select(daily, longitude,
                          latitude, location))
```

```{r}
sarajevo_map <- get_map(location = "Sarajevo Bosnia",
                        zoom = 12)


ggmap(sarajevo_map) +
  theme_map() +
  geom_point(data = map_data,
             aes(longitude,
                 latitude),
             size = 4) +
  geom_label_repel(data = map_data,
                   aes(longitude,
                       latitude,
                       label = location),
                   size = 10)
```

## WHO daily limit, PM10 in Sarajevo

```{r}
ggplot(daily) +
  geom_line(aes(day, pm10)) +
  facet_grid(location  ~ .) +
  geom_hline(yintercept = 50, col = "red") +
  ylab(expression(paste("Daily mean PM10 concentration (", mu, "g/",m^3,")"))) + 
  xlab("Time (days)") +
  theme(strip.text.y = element_text(angle = 0))
```

## Coyhaique vs. Ulaanbaatar

## 4th of July in the USA

## Data surfing

# What can you do?

## How to learn R

* AQ as a motivator like baseball?

## How to contribute to ropenaq

* Help build the documentation.

* Show me examples of figures, for adding how-to in the documentation.